\plaintitle

### 2016-05-28 cscw/papers/2017/labor/trackingWork.tex
inexplicably, we can keep submitting, so i added some citations i had really wanted to add

\begin{abstract}
  With growing attention in on--demand labor
  --- ranging from the ``sharing economy'' to information work ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  predicts and indeed precipitates both the contributions
  that have been made toward the development of on--demand labor and crowd work
  as well as the fallout among workers and researchers with regard to the disillusionment and alienation of work.
  % not only suggests, but in fact generates what we have observed:
  % members of this transient workforce increasingly feeling
  % disempowered,
  % marginalized,
  % and frustrated by the systems and platforms on which they work.

  After evaluating this framing
  through a series of case studies,
  % in various on--demand industries,
  we turn our lens to look to the future,
  to identify worthwhile questions and
  points of inquiry that researchers in social computing should consider
  as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

@@

\section{Introduction}
% \msb{I think a contribution of this paper can be, and should be, unifiying crowd work, sharing economy, and gig economy all under one theoretical banner: piecework. Come out and say that super visibly somewhere in this intro as a major contribution.}

The past decade has seen
paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has made connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work,
but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
or crowd work itself,
has proven difficult;
% as a result,
the connection between crowd work and a historical parallel contextualizing its ongoing developments 
hasn't been deeply made yet
\cite{turkopticon,dynamo,uberAlgorithm}.
% \msb{A weak differentiating sentence: }
% But the connections between
% historical piecework and
% contemporary crowd work
% haven't been deeply made as yet.
% um, note to self: say more.
% \msb{This paragraph doesn't set up the tension and release effectively. I suggest focusing this paragraph on a brief summary of the things that people have pointed out (low pay, Taylorism, difficult collective action). Basically paint them as a bunch of disconnected observations --- which is what they have been. Then at the start of the next paragraph, come out with a clear and visible thesis: say that all of these things fall under the banner of piecework. (and now you can finally mention that while piecework has been brought up, nobody has made the historical connections yet) Then you can move on to the next paragraph below, which is about how it's basically a necessary outcome.}

@@

\subsubsection{Low Pay}\label{sec:lowPay}
% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}

@@

This topic can be condensed into three major areas:
\begin{inlinelist}
\item external (scientific) management, and the evaluation of workers as functional modules; and
% \item the alienation of workers under this form of evaluation; and
\item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
\end{inlinelist}

@@

The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
and even physiological study of pieceworkers closely resembles the
research into cognitive loads and stress levels that we discussed among on--demand crowd workers
\cite{pieceworkBiologicalHarm,hu1961parallel}.
Even the complicated relationships between observers and workers themselves are not necessarily new
\cite{riisOtherSideLives}.
\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined among Turkers
\cite{facesOfMechanicalTurk}.
We argue that these investigations naturally follow the flourishing and fallout of what is,
fundamentally, piecework mediated in new ways.

% \subsubsection{Alienated Workers}
% Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
% T

% \itshape

% Marx's big entrance. Lily has lots of work pointing to this, and
% lots of that work points to yet more work gesturing in this direction.
% I want to bring a lot of that work
% that hasn't made it into the discussion yet
% into the conversation here.

@@

\subsubsection{Resistance}
It shouldn't surprise us, then, that workers have resisted the management imposed on them both by people and by systems,
often without recourse or room for feedback.
Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
many toggle their availability to avoid being dispatched to more distant locations,
resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
\cite{uberAlgorithm}.

@@

For insight on this, we return to \citeauthor{levi2009union}'s study of labor unions,
and identify that
``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
\cite{levi2009union}.
This perception already appears to be emerging in digitally mediated peer--governed organizations
\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
If these organizations and others are to avoid the same fate that labor unions faced,
they should take care to study this phenomenon and attempt to avoid it.

@@

Without claiming to have easy, cut--and--dry solutions to these problems,
we can nevertheless bring to attention a number of critical opportunities to
learn from historical parallels in piecework and factory labor,
and make informed decisions regarding whether
(or indeed how)
we may want to influence outcomes.
The challenges we bring to attention here are as follows:
\begin{inlinelist}
  \item codifying investment toward collective goods into the designs of systems;
  \item (re--)decentralizing the internet (hey, it's lowercase now!); and
  \item enabling reputation transferral.
\end{inlinelist}

@@

% \rule{\linewidth}{1pt}

% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case

% \textit{I've been going back and forth regarding how to frame this paper;
% the approach that tries to look at
% things that are similar vs things that are different
% doesn't seem to work, but
% I've left the thoughts here because it's not all completely bad.}

% \section{Things stay the same}
% How is gig work the same as it's been historically?

% I think this section would be compelling to draw parallels between
% the narratives drivers gave about the flexibility, autonomy, etc\dots
% and that which we might have seen among pieceworkers
% (predominantly women, who benefited from being able to work from home).

% \subsection{Flexibility}
% Are there cultural differences between the people that did piecework and
% the people that do gig work now?
% I'm not sure there are significant differences that have affected the outcomes so far.

% Many of the workers to whom piecework appealed were mothers, wives, etc\dots
% who mostly stayed at home for various reasons
% (certainly largely it was cultural
% --- women weren't afforded equal access to labor opportunities,
% making in--home job opportunities not only compelling, but also one of few available options).

% Gig workers are in some senses similarly constrained:
% workers on Amazon Mechanical Turk
% --- those that use it as a primary source of income, at least ---
% report being homebound for various reasons (e.g. medical, parenting, etc\dots).
% Society and circumstance have made it difficult
% or impossible
% to join the contemporary, conventional workforce;
% gig work re--opens that door.

% But there are differences; during our research over the summer,
% we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
% Many of them told us about their home lives
% --- about children, spouses, and other commitments ---
% to which they wanted to dedicate more time.
% One driver
% (let's call him Ra\'{u}l)
% told me about how he drove for Lyft
% after working as an inventory manager at a hospital for more than nine years.

% I asked him why he quit that job and forewent
% the benefits,
% predictability,
% and career growth opportunity
% that his old job offered.
% He told me that when his daughter was born,
% he was overcome with a desire to spend more time with her.
% No longer satisfied with work
% where he often left before his daughter would wake up and return after she fell asleep,
% Ra\'{u}l decided to start driving for Lyft,
% because he could drive in the evenings when his daughter was asleep.

% Other drivers reported similar benefits;
% gig work affords its workers flexibility that conventional careers don't allow.

% \subsection{routinization of work}
% This leads into the next section, but I want to bring up the process of
% making work about mass--manufacturing, at least inasmuch as
% the instructions are the same for everyone.

% The Internet arguably has made it much easier to broadcast those instructions,
% but it's had this deeper effect of enabling some amount of back--and--forth
% between the worker and the (algorithmic) manager.

% \subsection{Taylorism}
% The routinization of work makes it possible to measure that process,
% optimize it for certain characteristics, and ultimately
% lead to Taylorism and scientific management.
% This is not new;
% researchers have studied and written about the slow creep of
% algorithmic management and 
% discretization \& routinization of work tasks.

% \textit{We hope to take a step back from the context in which this work is often applied, and
% look for its place in the larger trends and theories
% to make sense of the trends of gig work at large.}

% Industrialization and the automobile assembly line makes this famous,
% but piecework functioned on the principle that everyone was making similar
% or identical
% garments and other products.

% Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
% sometimes bootstrapped,
% as in Ranjay's talk on 
% ``Embracing Error to Enable Rapid Crowdsourcing'',
% but more conventionally in work flows like ``Find--Fix--Verify''.

% \textit{This might be an opportunity to reflect on
% how pieceworkers internalized the work they were doing,
% responded to the stress of the uncertainty of potentially rejected work,
% etc\dots but I'm not familiar with research in that space.}

% \subsection{The emergence of decentralized workplaces}
% The practice of in--home piecework was consumed by
% the centralization of factories
% (the effect of which we'll talk about later, since
% we can talk about how this made unions more practical),
% but for a time many pieceworkers
% at the turn of the \nth{20} century
% worked out of their homes.
% Strikingly, many of the gig workers
% \citeauthor{crowdworkFuture}
% discussed in
% \citeyear{crowdworkFuture},
% and indeed many more continue to do what we might call ``information work''
% --- that is, work that predominantly demands human computation ---
% but increasingly we're seeing the movement toward transient work that largely requires embodied presence
% \cite{uberAlgorithm} (and others).

% \subsection{What was this about?}
% I was going to check in quickly and take a photo of your whiteboard
% (hoping that you'd kept our conversation notes around)
% but your room was \textit{literally} full of people lol.

% \section{Things are changing}
% The medium on which this work is being done
% --- and to an extent the medium used to manage workers ---
% has dramatically changed things as well, however;
% workers are distributed around the world,
% working out of their cars in the cases of livery services
% (notably, never returning to a base of operations)
% across and between cities as well as nations,
% or in their homes
% (paralleling the trend of piecework even more closely).

% Trying to understand how gig work has differed from piecework should at least
% start with looking at the different characteristics of the work involved.
% After that, we should think about how
% the demography and culture of the people engaging in this kind of work
% have changed versus that of the pieceworkers.

% \subsection{Differences in the work itself}
% Gig work has all of the above similarities with piecework,
% but there are key differences.

% % Piecework emerged at the turn of the \nth{20} century,
% % right at the time that telecommunications began to boom in the United States. % too far a reach?
% % \textit{Did telecom enable remote management?}

% Gig work in its contemporary formation is largely mediated by
% ubiquitously accessible digital media
% (the Internet, telephony, etc\dots)
% and importantly has relied on this technology to facilitate the remote management of workers
% \cite{uberAlgorithm}.

% \textit{Has the work fundamentally changed,
% or are we just being managed remotely in different ways?
% I'm not entirely sure.}

% How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
% How does this work differ from the work in which truckers, taxi drivers, and other
% independent contractors have been participating for decades?
% We argue that the substantive difference in these markets is the speed of the market itself,
% motivated by the technology which mediates it.
% Because workers can be sourced and dispatched virtually instantly,
% businesses that engage in this kind of work
% (e.g. Uber, Amazon Mechanical Turk, etc\dots)
% have taken to removing other bottlenecks,
% like vetting workers upfront

% \section{More deliberate work}
% \subsection{Differences in the people that do the work}
% ???

>>>>

\begin{abstract}
  With growing attention toward on--demand labor
  --- ranging from the ``sharing economy'' to information work ---
  scholars have made connections to various frameworks 
  and mechanisms such as worker advocacy, empowerment, and Taylorism,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that the literature surrounding ``piecework''
  informs and even predicts both the contributions
  that have been made toward the development of on--demand labor and crowd work
  as well as the fallout among workers and researchers
  with regard to the disillusionment and alienation of work.

  After evaluating this framing
  through a series of case studies,
  we look to the future
  to identify worthwhile questions and
  points of inquiry, such as the movement toward factories,
  that researchers in social computing should consider
  as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

@@

\section{Introduction}
The past decade has seen
paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has made connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work,
but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
or crowd work itself,
has proven difficult;
the connection between crowd work and a historical parallel contextualizing its ongoing developments 
hasn't been deeply made yet
\cite{turkopticon,dynamo,uberAlgorithm}.

@@

\subsubsection{Low Pay}\label{sec:lowPay}

@@

This topic can be condensed into two major areas:
\begin{inlinelist}
\item external (scientific) management, and the evaluation of workers as functional modules; and
% \item the alienation of workers under this form of evaluation; and
\item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
\end{inlinelist}

@@

The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
and even physiological study of pieceworkers closely resembles the
research into cognitive loads and stress levels that we discussed among on--demand crowd workers
\cite{pieceworkBiologicalHarm,hu1961parallel}.
Even the complicated relationships between observers and workers themselves are not necessarily new;
\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
\cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.

Similarly, \citeauthor{pollard1963factory}'s words on the punishment factory workers faced ---
for example, that ``unsatisfactory work was punished \dots~by fines or by dismissal'' ---
seems especially relevant given the fears we now know to be ubiquitous on platforms such as AMT, Uber, and other on--demand markets
\cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.

% \subsubsection{Alienated Workers}
% The dissatisfaction on crowd work systems is not as well--explored as their actual responses to that dissatisfaction
% (in various forms of resistance),
% but it's nevertheless worthwhile to point out some of the observations that have been made here,
% as external management, discussed earlier, and the discussion on worker resistance, to be discussed next,
% are bridged by this quality of on--demand labor markets.

% Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
% T

% \itshape
% Marx's big entrance. Lily has lots of work pointing to this, and
% lots of that work points to yet more work gesturing in this direction.
% I want to bring a lot of that work
% that hasn't made it into the discussion yet
% into the conversation here.

@@

\subsubsection{Resistance}
It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,
often without recourse or opportunity for feedback, let alone substantive input.
Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
many toggle their availability to avoid being dispatched to more distant locations,
resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
\cite{uberAlgorithm}.

@@

For insight on this, we return to \citeauthor{levi2009union}'s study of labor unions,
and identify that
``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
\cite{levi2009union}.
% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
% unable to satisfy their needs or otherwise more invested in the 
This perception already appears to be emerging in digitally mediated peer--governed organizations,
as \citeauthor{keegan2010egalitarians} and others have illustratively documented
\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
If these organizations and others are to avoid the same fate that labor unions faced,
they should take care to study this phenomenon and attempt to avoid it.

@@

Without claiming to have easy, cut--and--dry solutions to these problems,
we can nevertheless bring to attention a number of critical opportunities to
learn from historical parallels in piecework and factory labor,
and make informed decisions regarding whether
(or indeed how)
we may want to influence outcomes.
The challenges we bring to attention here are as follows:
\begin{inlinelist}
  \item codifying investment toward collective goods into the designs of systems;
  \item (re--)decentralizing the internet; and
  \item enabling reputation transferral.
\end{inlinelist}

@@
